---
title: "Tables"
author: "Robin Evans"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This package contains methods for storing and manipulating collections of contingency 
tables, and for easily vectorizing functions which apply to a contingency table.  

The basis of this is the class of object `tables`, which contains a collection of 
numerical tables all of the same dimension.  Let's create a collection of 10 
contingency tables (in this case probability tables), each of dimension 2x2x2.
```{r}
library(contingency)
tab <- rprobMat(10, 2, 3)
tab
```
The print method shows the first table in the list.  
The tables are stored as a matrix as can be seen by using the `dim()` function.  Accessing
particular rows of this matrix return the appropriate tables:
```{r}
tab[c(1,4,5),]
```
However we can also specific elements of the tables using their co-ordinates, and 
(optionally) leaving the first entry blank:
```{r}
tab[,1,1,]
```
Some basic operations are predefined, such as taking the margin of each table, or 
calculating a conditional distribution.
```{r}
margin(tab, 2:3)         # margin of second and third dimensions
conditional(tab, 2, 1)  # second dimension conditional on first
```

## More...

